Anya Jacoby
---
title: "p8105_hw1_abj2136"
output: github_document
date: "2026-09-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

### Loading packages and data set for the problem:
```{r packagesm, message = FALSE}
library(tidyverse)
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset contains `r nrow(penguins)` observations and `r ncol(penguins)` variables, including `r (names(penguins))`. The data set includes penguins from three islands (`r unique(penguins$island)`) and three unique species (`r unique(penguins$species)`). The data was collected from `r min(penguins$year)` to `r max(penguins$year)` and contains information on bill length, bill depth, flipper length, body mass and sex of the penguins.

The mean flipper length among the penguins in this data set is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE, 2))` mm. 

### Making plots:
```{r penguin_scatterplot, message = FALSE, warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("penguin_q1_scatterplot.png")
```
# Problem 2

### Creating dataframes:

```{r q2_dataframe, message = FALSE}

# setting seed for random sample
set.seed(1)

#creating dataframe
penguin_df = tibble (
  norm_samp = rnorm(10),
  greater_than_zero = norm_samp > 0,
  char_vec = letters[1:10],
  factor_vec = factor(sample(c(2,4,6), 10, replace = TRUE))
)

```

### Attempting to calculate means:

```{r q2_means}

penguin_df |> pull(norm_samp) |>  mean()
penguin_df |> pull(greater_than_zero) |> mean()
penguin_df |> pull(char_vec) |> mean()
penguin_df |> pull(factor_vec) |> mean()

```

It is possible to calculate the mean of the random sample (norm_samp) and the logical vector (greater_than_zero) because these are stored as numeric. The random sample contains 10 random numbers that can be averaged, and the logical vector contains a series TRUE or FALSE that is converted to 1's and 0's that can be averaged. 

It is not possible to calculate the mean of the character vector (char_vec) because this is stored as a character without a numeric value. Similarly, even though the factor vector (factor_vec) uses numbers as the levels, these are treated as distinct categorical labels, not numberic values, so a mean can not be calculated. That is why warning messages appear for these two lines of code stating that the arguments are not numeric or logical.   

### Converting to numeric values
```{r q2_convert_to_numeric, results = 'hide'}

penguin_df |> pull(greater_than_zero) |> as.numeric()
penguin_df |> pull(char_vec) |> as.numeric()
penguin_df |> pull(factor_vec) |> as.numeric()

```
The logical vector is converted to a series of 1's and 0's, the numeric values for TRUE and FALSE. The character vector is turned into a series of NA's, which causes a warning because character strings can not be converted into numbers, they do not have set numeric values (I kept the warning in to show this). The factor vector is converted into numbers based on the rank of each inputted value's level, but not based on the actual numeric value, which is why the level values (1,2,3) differ from the values entered into the factor vector (2,4,6).  
