Anya Jacoby
---
title: "p8105_hw1_abj2136"
output: github_document
date: "2026-09-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

### Loading packages and data set for the problem:
```{r packagesm, message = FALSE}
library(tidyverse)
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset contains `r nrow(penguins)` observations and `r ncol(penguins)` variables, including `r (names(penguins))`. The data set includes penguins from three islands (`r unique(penguins$island)`) and three unique species (`r unique(penguins$species)`). The data was collected from `r min(penguins$year)` to `r max(penguins$year)` and contains information on bill length, bill depth, flipper length, body mass and sex of the penguins.

The mean flipper length among the penguins in this data set is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE, 2))` mm. 

### Making plots:
```{r scatterplot, message = FALSE, warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("penguin_q1_scatterplot.png")
```
# Problem 2

### Creating dataframes:

```{r}

set.seed(1)

penguin_df = tibble (
  norm_samp = rnorm(10),
  greater_than_zero = norm_samp > 0,
  char_vec = letters[1:10],
  factor_vec = factor(sample(c(1,2,3), 10, replace = TRUE))
)



```



